---
title: "BasicAnalysis"
author: "martinvanelp"
date: "Sunday, September 21, 2014"
output: pdf_document
---

##Basic inferential data analysis
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 
1. Load the ToothGrowth data and perform some basic exploratory data analyses.
2. Provide a basic summary of the data.
3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. (Use the techniques from class even if there's other approaches worth considering)
4. State your conclusions and the assumptions needed for your conclusions.

##Output
**1. Load the ToothGrowth data and perform some basic exploratory data analyses.**

**2. Provide a basic summary of the data.**
```{r}
library(datasets)
summary(ToothGrowth)
str(ToothGrowth)
with(ToothGrowth, { 
        par(mfrow = c(1,3))
        hist(len)
        hist(dose)
        plot(len, dose)
})
```

There are three variables in the dataset: two numeric (len & dose) and one factor (supp). Variable len seems sort of normally distributed. It seems dose and length are correlated given the last plot.

**3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose.** (Use the techniques from class even if there's other approaches worth considering)

On a 95% confidence interval, the supplement is not significantly influencing tooth growth in this sample. Each increase in dose increases the tooth growth significantly on a 95% confidence interval in this sample. (see Appendix for the underlying t-tests)

**4. State your conclusions and the assumptions needed for your conclusions.**

For tooth growth the dose of the supplement significantly influential. The kind of supplement on the other hand does not significantly influence tooth growth, given a 95% confidence interval in both cases. The underlying assumption is that the sample is representative for the population.

##Appendix
```{r}
TGsupp <- subset(ToothGrowth, supp %in% c("OJ", "VC"))
t.test(len ~ supp, paired = FALSE, var.equal = TRUE, data = TGsupp)

TG12 <- subset(ToothGrowth, dose %in% c(0.5, 1.0))
t.test(len ~ dose, paired = FALSE, var.equal = TRUE, data = TG12)

TG13 <- subset(ToothGrowth, dose %in% c(0.5, 2.0))
t.test(len ~ dose, paired = FALSE, var.equal = TRUE, data = TG13)

TG23 <- subset(ToothGrowth, dose %in% c(1.0, 2.0))
t.test(len ~ dose, paired = FALSE, var.equal = TRUE, data = TG23)
```
